### 2017-04-05 paper/oopsla17-submission/learning.tex
more learning algos

There are many learning algorithms to choose from, existing
on a spectrum that balances expressiveness with ease of training (and of
interpreting the learned model).
%
In this section we present four standard learning algorithms that we
will investigate:
%
(1) logistic regression,
(2) decision trees,
(3) random forests, and
(4) neural networks.
%
For each algorithm, we describe the form of the model it learns, the
training procedure, and any hyper-parameters that must be supplied by a
human.

@@

The simplest model we investigate is logistic regression, which can be
understood in two steps.
%
First, we compute the \emph{evidence} for a label $y$ as a linear
function of the feature vector $\mathbf{x}$, using a weight vector
$\mathbf{w}$ and a bias $b$ that will be learned from the training set.
$$
\mathsf{ev} = \sum_j w_j x_j + b
$$
Then we use the logistic function (also called a sigmoid function)
$\sigma(x) = 1 / (1 + e^{-x})$ to compress the evidence term into the
$[0,1]$ interval, which can then be interpreted as the probability that
the label $y$ should be applied.
%
This gives us the final model:
$$
Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + \mathsf{exp}(-\sum_j w_j x_j - b)}
$$

@@

\paragraph{Decision Trees}
Decision tree algorithms learn a tree of binary predicates over the
features, recursively partitioning the input space until a final
classification can be made.
%
Each node in the tree contains a single predicate of the form
$x_j \leq t$ for some threshold $t$, which determines whether a given
input should proceed down the left or right subtree.
%
Each leaf is labeled with a prediction and the fraction of training
samples that would reach it, the latter quantity can be interpreted as
the decision tree's confidence in its prediction.
\ES{HUMA: is this correct?}

@@

\paragraph{Random Forests}
\ES{TODO: Maybe Huma?}

\paragraph{Neural Networks}
The last (and most complex) model we discuss is a type of neural network
called a \emph{multi-layer perceptron} \ES{CITE}.
%
Neural networks can be understood as a collection of linear models, just
like the one used in logistic regression, arranged in a directed graph.
%
Each node in the graph corresponds to a single linear model, called a
\emph{neuron}, and the edges propagate signals from the input features,
through the neurons, to the output labels.
%
The linear model in each neuron is wrapped in an \emph{activation
  function}, which controls whether (or how strongly) the signal will be
propagated out.
%
These activation functions are what allow neural networks to accurately
model nonlinear functions.
%
A common choice of activation function is the same sigmoid function
used by logistic regression, alternatives include the hyperbolic $\tanh$
function, and more recently the \emph{rectified linear unit} (ReLU)
$f(x) = \mathsf{max}(0,x)$ \ES{CITE}.

@@

Neural networks are trained similarly to logistic regression using
stochastic gradient descent, and thus the learning rate $\eta$ and
regularization rate $\lambda$ are also important hyper-parameters.

>>>>

There are many learning algorithms to choose from, existing
on a spectrum that balances expressiveness with ease of training (and of
interpreting the learned model).
%
In this section we present four standard learning algorithms that we
will investigate:
%
(1) logistic regression,
(2) decision trees,
(3) random forests, and
(4) neural networks.
%
For each algorithm, we describe the model it learns, the training
procedure, any hyper-parameters that must be tuned by a human, and a
summary of its advantages and disadvantages.

@@

The simplest model we investigate is logistic regression, which can be
understood in two steps.
%
First, we compute the \emph{evidence} for a label $y$ as a linear
function of the feature vector $\mathbf{x}$, using a weight vector
$\mathbf{w}$ and a bias $b$ that will be learned from the training set.
$$
\mathsf{ev} = \sum_j w_j x_j + b
$$
Then we use the logistic function (also called a sigmoid function)
$\sigma(x) = 1 / (1 + e^{-x})$ to compress the evidence term into the
$[0,1]$ interval, which can then be interpreted as the probability that
the label $y$ should be applied.
%
This gives us the final model:
$$
Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + \mathsf{exp}(-\sum_j w_j x_j - b)}
$$
% $$
% Pr(y\ |\ \mathbf{x}) = \frac{1}{1 + e^{-\mathsf{ev}}}
% $$

@@

\paragraph{Decision Trees}
Decision tree algorithms learn a tree of binary predicates over the
features, recursively partitioning the input space until a final
classification can be made.
%
Each node in the tree contains a single predicate of the form
$x_j \leq t$ for some feature $x_j$ and threshold $t$, which determines
whether a given input should proceed down the left or right subtree.
%
Each leaf is labeled with a prediction and the fraction of training
samples that would reach it, the latter quantity can be interpreted as
the decision tree's confidence in its prediction.
\ES{HUMA: is this correct?}

@@

\paragraph{Random Forests}
One technique for increasing the accuracy of a classifier is to train an
\emph{ensemble} of distinct classifiers and use a majority vote
to make the final prediction \ES{CITE?}.
%
Random forests \ES{CITE} are an instance of this method, with decision
trees as the underlying classifiers.
%
The training process involves taking $N$ random subsets of the training
data and training a separate decision tree on each subset --- the
training process for the decision trees is often modified slightly to
reduce correlation between trees, by forcing the tree to pick features
from a random subset of all features at each node.
%
In order to make a prediction for an unseen sample, the random forest
asks each decision tree to make a prediction, and then predicts the most
common label.

The diversity of the underlying models tends to make random forests less
susceptible to the overfitting and instability of decision trees, but it
also makes the learned model more difficult to interpret.
%
Furthermore, the number of trees in the forest, $N$, becomes an important
hyper-parameter that must be set manually.

\paragraph{Neural Networks}
The last (and most complex) model we discuss is a type of neural network
called a \emph{multi-layer perceptron} \ES{CITE}.
%
Neural networks can be understood as a collection of linear models, just
like the one used in logistic regression, arranged in a directed graph.
%
Each node in the graph corresponds to a single linear model, called a
\emph{neuron}, and the edges propagate signals from the input features,
through the neurons, to the output labels.
%
The linear model in each neuron is wrapped in an \emph{activation
  function}, which controls whether (or how strongly) the signal will be
propagated out.
%
These activation functions are what allow neural networks to accurately
model nonlinear functions.
%
A common choice of activation function is the same sigmoid function used
by logistic regression, alternatives include the hyperbolic $\tanh$
function, and more recently the \emph{rectified linear unit} (ReLU),
defined as $f(x) = \mathsf{max}(0,x)$ \ES{CITE}.

@@

Neural networks are trained similarly to logistic regression using
stochastic gradient descent, and thus the learning rate $\eta$ and
regularization rate $\lambda$ are also important hyper-parameters.
%
They are, however, much more difficult to train as they require larger
training sets.

Neural networks are popular for their ability to uncover ``hidden''
features in the data.
%
Each neuron learns a different linear combination of the signals from
the previous layer; combined with the activation function this can be
thought of as the neuron learning to respond to a new, \emph{compound}
feature of its inputs.
%
This ability to discover interesting combinations of features makes
neural networks easier to use in some respect; they can reduce the
importance of manual \emph{feature engineering} (selecting interesting
features to provide as inputs).
%
However, the sheer size and complexity of neural networks makes the
model particularly difficult to interpret --- they are generally regarded
as ``black boxes.''
%
Neural networks are also very susceptible to overfitting, so
regularization methods like the $L_2$ norm used for logistic regression
are important.
